The complex structure and peculiar internal motions of the planetary nebula NGC 6210

### 2020-05-23 ms.tex
More on the outflow axes

    r % Axis

@@

    NE Blue complex & 60 &  & -51 \pm 1 & 23 \pm 5 & 65 \pm 2 & -69 \pm 4 & 5 \pm 1 & 0.14 \pm 0.04 & 2.1 \pm 0.6\\

@@

    SW Red complex & 250 &  & +38 \pm 1 & 18 \pm 3 & 49 \pm 2 & 68 \pm 3 & 7 \pm 2 & 0.19 \pm 0.06 & 3.7 \pm 1.2\\
    N(E) Red complex & 30 &  & +25 \pm 2 & 5 \pm 4 & 30 \pm 2 & 81 \pm 7 & 4 \pm 1 & 0.26 \pm 0.21 & 8.0 \pm 6.7\\
    SW Faint Blue & 225 &  & -26 \pm 1 & 26 \pm 4 & 41 \pm 3 & -50 \pm 4 & 7 \pm 1 & 0.11 \pm 0.02 & 2.6 \pm 0.5\\

@@

    NW knot  & 310 &  & +18 \pm 1 & 11 \pm 8 & 24 \pm 4 & 63 \pm 17 & 8 \pm 1 & 0.18 \pm 0.10 & 6.7 \pm 4.5\\
    N(W) Red complex & 300 &  & +21 \pm 2 & 10 \pm 4 & 27 \pm 3 & 68 \pm 8 & 6 \pm 2 & 0.16 \pm 0.08 & 6.0 \pm 3.1\\

@@

    SE Blue complex & 135 &  & -31 \pm 3 & 39 \pm 10 & 54 \pm 8 & -44 \pm 8 & 5 \pm 1 & 0.07 \pm 0.02 & 1.1 \pm 0.4\\
    SE knot & 135 &  & -35 \pm 3 & 39 \pm 10 & 57 \pm 7 & -47 \pm 8 & 8 \pm 1 & 0.09 \pm 0.02 & 1.5 \pm 0.4\\

@@

    (3)~Outflow axis that feature is best aligned with, if any.

@@

For those nebular features where both proper motion and radial velocity measurements exist,
it is possible to estimate the full three-dimensional velocity vector.
Under the assumption that all motions are strictly radial,
one can also find the three-dimensional position.
These are given in Table~\ref{tab:3d} for a wide variety of nebular features
in the inner shells, intermediate shell, low-ionization knots, and outer lobes.\footnote{
  Note that the uncertainties given in the table are derived from observational errors only,
  and they do not include the systematic uncertainty in the distance.
}

When combining radial velocities and proper motion measurements,
it is necessary to account for the fact that radial velocities measure the material speed,
whereas proper motions measure pattern speeds of shocks or ionization fronts \citep{Mellema:2004a}. 
The exact correction factor (material speed divided by pattern speed)
is model-dependent (e.g., Appendix~A of \citealp{ODell:2009c}),
but is predicted to be generally larger than unity
\citetext{see Fig.~8 of \citealp{Jacob:2013a}}
unless the nebula is in the recombination phase.
For simplicity, we adopt a constant correction factor of 1.2 for all entries in Table~\ref{tab:3d},
which is consistent with the predicted value for expansion velocities greater than \SI{30}{km.s^{-1}}
\citetext{see Fig.~3 of \citealp{Schonberner:2019a}}.

@@

An approximate three-dimensional visualization of the nebular structure is given in Figure~\ref{fig:cut-axis-3d}.
The three axes, A, C, and D, that define the outer lobes all have small inclinations,
\(|i| < \ang{30}\) to the plane of the sky,
while the knot complexes all have higher inclinations,
\(|i| > \ang{45}\).

It is apparent from this figure that the three axes that define the outer lobes,
A, B, and D,
can be grouped into a loose conical structure with a half-opening angle of 
\begin{figure}
  \includegraphics[width=\linewidth]
  {figs/vel-radius-systems-annotated.pdf}
  \caption{
    True velocity, \(V_{\text{tot}}\), versus true radius, \(R\), for all the features listed in Table~\ref{tab:3d}.
    Diagonal lines show dynamical times between \num{1000} and \SI{16000}{years}, as marked.
    Symbol shape indicates the type of feature:
    circle symbols are high-ionization shells (\S~\ref{sec:high-ioniz-shells}),
    cross symbols are low-ionization knots or knot complexes (\S~\ref{sec:knot-complexes}),
    diamond symbols are high-velocity and intermediate-velocity systems in the outer lobes (\S~\ref{sec:outer-lobes}).
    Error bars show the observational uncertainties given in the table,
    except for outer lobe system \SysM{2}, where the lack of proper motion measurements mean that only lower limits for \(V_{\text{tot}}\) and \(R\) are available, indicated by arrows.
    The color of each symbol corresponds the position angle of the feature, as shown by the key at right,
    which also indicates the values for each of the four axes: A, B, C, and D.
    The position angle is calculated modulo \ang{180} so that the positive and negative arm of each axis have the same color.
    Also shown (gray ellipses) are the expansion velocity and range of radii of the inner and outer halo features (\S~\ref{sec:haloes}).
      }
  \label{fig:ages}
\end{figure}

@@

>>>>

    l % Axis

@@

    NE Blue complex & 60 & \AxM{E} & -51 \pm 1 & 23 \pm 5 & 65 \pm 2 & -69 \pm 4 & 5 \pm 1 & 0.14 \pm 0.04 & 2.1 \pm 0.6\\

@@

    SW Red complex & 250 & \AxP{E} & +38 \pm 1 & 18 \pm 3 & 49 \pm 2 & 68 \pm 3 & 7 \pm 2 & 0.19 \pm 0.06 & 3.7 \pm 1.2\\
    N(E) Red complex & 30 & \AxP{E} & +25 \pm 2 & 5 \pm 4 & 30 \pm 2 & 81 \pm 7 & 4 \pm 1 & 0.26 \pm 0.21 & 8.0 \pm 6.7\\
    SW Faint Blue & 225 & \AxM{E} & -26 \pm 1 & 26 \pm 4 & 41 \pm 3 & -50 \pm 4 & 7 \pm 1 & 0.11 \pm 0.02 & 2.6 \pm 0.5\\

@@

    NW knot  & 310 & \AxP{E} & +18 \pm 1 & 11 \pm 8 & 24 \pm 4 & 63 \pm 17 & 8 \pm 1 & 0.18 \pm 0.10 & 6.7 \pm 4.5\\
    N(W) Red complex & 300 & \AxP{E} & +21 \pm 2 & 10 \pm 4 & 27 \pm 3 & 68 \pm 8 & 6 \pm 2 & 0.16 \pm 0.08 & 6.0 \pm 3.1\\

@@

    SE Blue complex & 135 &  \AxM{E} & -31 \pm 3 & 39 \pm 10 & 54 \pm 8 & -44 \pm 8 & 5 \pm 1 & 0.07 \pm 0.02 & 1.1 \pm 0.4\\
    SE knot & 135 & \AxM{E} & -35 \pm 3 & 39 \pm 10 & 57 \pm 7 & -47 \pm 8 & 8 \pm 1 & 0.09 \pm 0.02 & 1.5 \pm 0.4\\

@@

    (3)~Outflow axis that feature is best aligned with.

@@

For those nebular features where both proper motion and radial velocity measurements exist,
it is possible to estimate the full three-dimensional velocity vector.
Under the assumption that all motions are strictly radial,
one can also find the three-dimensional position.
These are given in Table~\ref{tab:3d} for a wide variety of nebular features
in the inner shells, intermediate shell, low-ionization knots, and outer lobes.\footnote{
  Note that the uncertainties given in the table are derived from observational errors only,
  and they do not include the systematic uncertainty in the distance.
}
When combining radial velocities and proper motion measurements,
it is necessary to account for the fact that radial velocities measure the material speed,
whereas proper motions measure pattern speeds of shocks or ionization fronts \citep{Mellema:2004a}. 
The exact correction factor (material speed divided by pattern speed)
is model-dependent (e.g., Appendix~A of \citealp{ODell:2009c}),
but is predicted to be generally larger than unity
\citetext{see Fig.~8 of \citealp{Jacob:2013a}}
unless the nebula is in the recombination phase.
For simplicity, we adopt a constant correction factor of 1.2 for all entries in Table~\ref{tab:3d},
which is consistent with the predicted value for expansion velocities greater than \SI{30}{km.s^{-1}}
\citetext{see Fig.~3 of \citealp{Schonberner:2019a}}.

@@

An approximate three-dimensional visualization of the nebular structure is given in Figure~\ref{fig:cut-axis-3d}.
The high-ionization shells have two distinct symmetry axes, A and B, which are close to orthogonal.
Axis~A is shared with the outer lobes but the lobes also present outflows along two additional axes, C and~D.
The low-ionization knot complexes and individual knots do not share any of these four axes,
with the notable exception of the N~knot, which lies along axis~A.

Despite the apparent complexity of the multi-axis structure, there are some clear systematic patterns,
one of which is illustrated in Figure~\ref{fig:inclinations}.
This shows histograms of the outflow inclination angle from the plane of the sky (absolute value)
for all of the individual features of Table~\ref{tab:3d}, classified by nebular component.
The inclination distribution for the outer lobes is very different from that of the knot complexes.
The lobes are concentrated at low inclinations, \(|i| = \ang{20 \pm 3}\),
whereas the knot complexes are concentrated at high inclinations, \(|i| = \ang{63 \pm 6}\).
The shells show intermediate inclinations, \(|i| = \ang{32 \pm 3}\),
while the individual knots are diverse
(the SE and NW~knot have high inclinations like the knot complexes,
while the N~knot has a low inclination like the lobes).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/turtle-inclination-histogram}
  \caption{Distribution of outflow inclinations for the different components.}
  \label{fig:inclinations}
\end{figure}

The concentration of high inclinations for the knot complexes imply the presence of a fifth axis in the nebula, which we denote axis~E.
Since this axis is close to the plane of the sky,
its signature is not obvious on images of the nebula but can only be discerned from spectroscopy.
If we take the average vector outflow direction of the knot complexes,
then axis~E has an inclination \(i = \ang{69}\)
and a position angle \(\mathrm{PA} = \ang{331}\).
The individual knot complexes have outflows that deviate by only \ang{16} to \ang{29} from this average direction,
confirming that it is indeed a well-defined axis,
albeit with a substructure of three point symmetric pairs of outflows,
as discussed in \S~\ref{sec:knot-complexes}.

\begin{figure}
  \includegraphics[width=\linewidth]
  {figs/vel-radius-systems-annotated}
  \caption{
    True velocity, \(V_{\text{tot}}\), versus true radius, \(R\), for all the features listed in Table~\ref{tab:3d}.
    Diagonal lines show dynamical times between \num{1000} and \SI{16000}{years}, as marked.
    Symbol shape indicates the type of feature:
    circle symbols are high-ionization shells (\S~\ref{sec:high-ioniz-shells}),
    cross symbols are low-ionization knots or knot complexes (\S~\ref{sec:knot-complexes}),
    diamond symbols are high-velocity and intermediate-velocity systems in the outer lobes (\S~\ref{sec:outer-lobes}).
    Error bars show the observational uncertainties given in the table,
    except for outer lobe system \SysM{2}, where the lack of proper motion measurements mean that only lower limits for \(V_{\text{tot}}\) and \(R\) are available, indicated by arrows.
    The color of each symbol corresponds the position angle of the feature, as shown by the key at right,
    which also indicates the values for each of the four axes: A, B, C, and D.
    The position angle is calculated modulo \ang{180} so that the positive and negative arm of each axis have the same color.
    Also shown (gray ellipses) are the expansion velocity and range of radii of the inner and outer halo features (\S~\ref{sec:haloes}).
      }
  \label{fig:ages}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]
  {figs/turtle-flow-axis-history}
  \caption{
    Ejection history of each flow axis, assuming no acceleration or deceleration of the features.
  }
  \label{fig:axis-history}
\end{figure}

\textbf{TODO Discussion of C and D being older than axis A}

@@

Velocity mapping of the Turtle Nebula, NGC~6210, in multiple emission lines
reveals a complex structure of shells and collimated outflows,
expanding along at least 5 different axes.

\begin{enumerate}[1.]
\item 
\item High-ionization 
\end{enumerate}
